---
title: "gapnider_dplyr"
author: "Rodney"
date: "19 June 2019"
output: html_document
---

you hit ctl_alt+I and you get the ``` sutff with the r below
```{r}

library(tidyverse)
install.packages(dplyr)
library(dplyr)

```

load data
```{r}
## loading the gapmider dataset
gapminder <- read.csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder.csv")

```
```{r}
head(gapminder)
tail(gapminder)
## str() - stands for strcuture

str(gapminder)

```

look at individual columns
```{r}
##show the lifeexp column
gapminder$lifeExp



```


Learning the 'filter()' function
```{r}
filter(gapminder, lifeExp < 35)

# just look at mexico
filter(gapminder, country == 'Mexico')
```

filter(gapminder, lifeExp < 29)

Average life expectaNCY IN BRAZIL
```{r}
brazil_life_exp <- filter(gapminder, country == 'Brazil', year > 1986, year < 2008)

brazil_life_exp
```

Learning the select 'select()' function

```{r}
select(gapminder, year, country, lifeExp)
select(gapminder, -continent, -gdpPercap, -pop) ## select removing variablke

```

Using 'filter()' and 'select()'

```{r}
TODO

```
using the pipe operator %>% 
cmd+shift+m
ctrl+shift+m

```{r}
gapminder %>% head()

gap_cambodia <- filter(gapminder, country =="Cambodia") %>%  # and then
select(-continent, -lifeExp)
gap_cambodia

TODO another combination 

```

Learn 'mutate()' which adds a new column

```{r}
gapminder %>% 
  mutate(gdp = pop * gdpPercap)

```
Ex 6.10.1

```{r}

gapminder %>% 
  filter(continent == "ASia", year = 2007) %>% 
  mutate(pop_thousands = pop/1000) %>% 
  select(country, year, pop_thousands) %>% 
  mean(pop_thousands)


```

Learning the group_by() function

```{r}
gapminder %>% 
  filter(year == 2002)
  group_by(continent) %>% 
  summarize(count_pop = sum (pop))
```

```{r}
df <- gapminder %>%  
  group_by(continent, year) %>% 
  summarize(cont_pop = sum(pop)) %>% 
  ungroup() ## good tip always to ungroup after you add a grouping variable
  
```


```{r}
df <- gapminder %>%  
  group_by(continent, year) %>% 
  summarize(cont_pop = sum(pop)) %>% 
  arrange(-year) ## good tip always to ungroup after you add a grouping variable
  
```

Ex

```{r}

##What is the maximum GDP per continent across all years?

df <- gapminder %>%
  mutate(gdp = pop * gdpPercap) %>% ## mutate create a new column gdp
  group_by(continent) %>%
  mutate(max_gdp = max(gdp)) %>%
  filter(gdp == max_gdp)

head(df)




```


What is the maximum GDP per continent across all years?

df <- gapminder %>%
  mutate(gdp = pop * gdpPercap) %>%
  group_by(continent) %>%
  mutate(max_gdp = max(gdp)) %>%
  filter(gdp == max_gdp)

head(df)










