
---
title: "Graphics with ggplot2"
author: "Rodney"
date: "11/21/2017"
output: html_document
---

# Learning ggplot2

We're learning ggplot2 It's going to be amazing. 

Load tidyverse
```{r tidyverse}
library(tidyverse)
```

Get the data from github

```{r data}

ohi_data <- read.csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/OHI_global_data.csv")

```

we start

```{r start ggplot}

ggplot(data=ohi_data, aes(x=georegion_one, y=OHI_score)) + geom_point()

```

ggplot(data=ohi_data, aes(x=georegion_one, y=OHI_score)) + geom_jitter(with=0.2)

# get barplot

```{r boxplot}
 ggplot(data=ohi_data, aes(x=georegion_one)) + geom_bar()

```

# get hitogram

```{r histogram}
TODO
```

Multiple geoms explorations

```{r multiple geoms}

ohi <-summary <-read.csv("https://github.com/OHI-Science/data-science-training/blob/master/data/OHI_scores_georegion_summary.csv")

ggplot(data=ohi_data, aes(y=OHI_score, x = HDI, color=georegion_one)) + geom_point()

ggplot(data =ohi_data) + geom(aes(y = )) TODO

```

adding a third variable

```{r}
ggplot(data=ohi_data, aes(x=OHI_score, y=HDI, color = coastal_pop)) + geom_point()

```

adding a third discrete variable

```{r}

?par


Color brewer
```{r}
install.packages("RColorBrewer")
library(RColorBrewer)
display.brewer.all()
my_palette <- brewer.pal(n=9, "YLorRd")

my_palette <- brewer.pal(n=12, "Set3")
ggplot(data=ohi_data, aes(x=OHI_score, y =HDI, color = georegion_one)) + 


